A Study of the Effectiveness of Usage Examples in REST API Documentation

### 2017-04-05 controlled_study.tex
back to the correct size + observation number

\subsubsection{Data Analysis}
We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples. We exported a table of data from the database behind web-based REST API explorer. For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body. The values for I1-4 used by the participants were manually coded to group the participant responses into categories. The categories help us determine the types of information that API client developers need in the API documentation to perform API tasks successfully. The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under  codes that were already applied. The first author applied the codes on the raw data artifact and provided a coding scheme to a co-author. The co-author applied the codes to 82/539 (10\% confidence interval with 95\% confidence level) API calls made by the participants. To resolve disagreements, the authors discussed and updated the codes accordingly.

@@

\textbf{Observation: Use of Data Type.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of \lstinline{String} instead of \lstinline{Integer}. To successfully complete T1, participants had to use an integer ID representing an \lstinline{Author} object given the author name. In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name. P1.12 mentioned the following: \textit{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}. The enhanced documentation showed one example of using numeric ID. 3 of the 10 participants from G2 ran into the same error. When faced with this error, participants used trial and error to eventually complete the task.

@@

\textbf{Observation: Use of Data Format.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples. Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API. The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array. For example: they tried with \lstinline{exclude=[1,4], exclude=1&4, exclude=4&&exclude=1, id!=1&id!=4} and other alternatives before eventually finding the right format as follows \lstinline{exclude=1,4}. Participant P1.6 mentioned the following: \textit{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''}. 7 out of the 16 participants in G1 faced this error. The enhanced documentation showed one example of using multiple IDs and all G2 participants could use the right format.

@@

\textbf{Observation: Use of Request Headers.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples. Tasks T5 and T6 required the use of a HTTP request header named \lstinline{Content-Type}. In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''. Experienced software engineers still had problem using this HTTP header as we found from the participants in G1. P1.12 mentioned the following: \textit{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters... Examples would really help - with a sample request and a sample response.''}  Even though G1 participants have years of experience with REST APIs, only one of them could use the \lstinline{Content-Type} header correctly without example. In contrast, 2 out of the 10 participants from G2 failed to use this header correctly as indicated in the example in the enhanced documentation. Moreover, 1 of the 2 G2 participants that didn't use the header first time, later used it correctly on a subsequent attempt.

@@

\textbf{Observation: Use of Interdependent API Calls.} We observed the G1 participants had problem completing API tasks that require multiple API calls without usage examples. Tasks T2 and T6 required the use of multiple API calls to complete successfully. Both tasks require an initial API call to find data using the API that is needed make a second API call. For example, P1.5 mentioned the following about T6, \textit{``The post I created is not visible and not editable, though I got 201. Not sure why''}. After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the \lstinline{publish} status. As a result, the post was not returned via the API for T6. P2.2 mentioned the following about T2: \textit{``In case of search posts by author, it's not clear how to find the author\_id. Although, I realized that users and authors are same in this case.''}

@@

\subsubsection{API Learning Obstacles}
Robillard et al. used a mixed approach of surveys and in-person interviews with professional software engineers to understand the API learning obstacles \cite{Robillard_what_makes} \cite{Robillard_a_field_study}. They found that many obstacles for API learnabilty were related to API documentation. They identified five impactful factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation. Uddin and Robillard surveyed 323 professional software engineers to understand the different ways in which a piece of documentation can be unfit for purpose \cite{g_uddin}. They identified ambiguity, incompleteness, and incorrectness as the three severest problems that lead API documentations to fail to answer the information needs of developers. Duala-Ekoko and Robillard performed a study with 20 programmers to understand the types of questions that developers have while using an unfamiliar API \cite{Duala-Ekoko:2012:AAQ:2337223.2337255}. They categorized the API related information needs into twenty generic questions that can be used to analyze the quality of API documentation. Myers et al. found that participants had limited success finding the relevant API elements to perform an API task using an enterprise API \cite{Myers_study}. Among other factors, they recommended providing code examples in the API documentation to help API client developers. Our work fits in the general space of studies of API documentation obstacles. We focus specifically on problems related to the absence of examples expanding within the area of REST APIs.

\subsubsection{Crowd-Sourced API Usage Examples}
Wang et al. performed an exploratory study to understand the current state of API related knowledge available on the Internet \cite{6462686}. They searched the web for usage examples related to the API of five popular Java libraries and found that on average API examples could be found for 77\% of the 4,637 APIs included in the selected libraries. Moreover, they found that the crowd-sourcing sites accounted for 93.7\% of the usage examples compared to 6.13\% that were published on the official documentation. An evaluation of Jadite showed that developers were three times faster to complete API tasks with access to auto-referred crowd-sourced usage examples~\cite{5295283}. Nasehi et al. analyzed StackOverflow threads to understand the characteristics of good and bad code examples based on user provided votes \cite{Nasehi_what_makes}. They found that commonly down-voted API related answers lacked code examples, explanations, and shortcoming of solutions. Treude et al. proposed a machine learning based approach to find relevant API documentation from StackOverflow using both the text and metadata found from the questions and answers~\cite{Treude:2016:AAD:2884781.2884800}. Kim et al. presented a technique to automatically augment code examples from the web using code search tools to Java APIs \cite{Kim:2009:AEJ:1747491.1747552}. Jiau et al. observed a severe inequality within the context of crowd-sourced API usage examples where most of the content were related to the popular APIs \cite{jiau}. Existing research on crowd-sourced API documentation has mostly focused finding usage examples of local APIs. In this paper, we focused on REST APIs, where the API client development programming language is agnostic to API implementation language. This distinction with local APIs makes it hard to leverage the techniques proposed in the aforementioned papers for finding crowd-sourced REST API usage examples.

>>>>

\subsubsection{Data Analysis}
We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples. We exported a table of data from the database behind web-based REST API explorer. For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body. The values for I1-4 used by the participants were manually coded to group the participant responses into categories. The categories help us determine the types of information that API client developers need in the API documentation to perform API tasks successfully. The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under  codes that were already applied. The first author applied the codes on the raw data artifact and provided a coding scheme to a co-author. The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level) made by the participants. The two authors validated the codes against each other to provide consistency.

@@

\textbf{Observation 1: Use of Data Type.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of \lstinline{String} instead of \lstinline{Integer}. To successfully complete T1, participants had to use an integer ID representing an \lstinline{Author} object given the author name. In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name. P1.12 mentioned the following: \textit{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}. The enhanced documentation showed one example of using numeric ID. 3 of the 10 participants from G2 ran into the same error. When faced with this error, participants used trial and error to eventually complete the task.

@@

\textbf{Observation 2: Use of Data Format.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples. Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API. The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array. For example: they tried with \lstinline{exclude=[1,4], exclude=1&4, exclude=4&&exclude=1, id!=1&id!=4} and other alternatives before eventually finding the right format as follows \lstinline{exclude=1,4}. Participant P1.6 mentioned the following: \textit{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''}. 7 out of the 16 participants in G1 faced this error. The enhanced documentation showed one example of using multiple IDs and all G2 participants could use the right format.

@@

\textbf{Observation 3: Use of Request Headers.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples. Tasks T5 and T6 required the use of a HTTP request header named \lstinline{Content-Type}. In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''. Experienced software engineers still had problem using this HTTP header as we found from the participants in G1. P1.12 mentioned the following: \textit{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters... Examples would really help - with a sample request and a sample response.''}  Even though G1 participants have years of experience with REST APIs, only one of them could use the \lstinline{Content-Type} header correctly without example. In contrast, 2 out of the 10 participants from G2 failed to use this header correctly as indicated in the example in the enhanced documentation. Moreover, 1 of the 2 G2 participants that didn't use the header first time, later used it correctly on a subsequent attempt.

@@

\textbf{Observation 4: Use of Interdependent API Calls.} We observed the G1 participants had problem completing API tasks that require multiple API calls without usage examples. Tasks T2 and T6 required the use of multiple API calls to complete successfully. Both tasks require an initial API call to find data using the API that is needed make a second API call. For example, P1.5 mentioned the following about T6, \textit{``The post I created is not visible and not editable, though I got 201. Not sure why''}. After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the \lstinline{publish} status. As a result, the post was not returned via the API for T6. P2.2 mentioned the following about T2: \textit{``In case of search posts by author, it's not clear how to find the author\_id. Although, I realized that users and authors are same in this case.''}

@@

\subsubsection{API Learning Obstacles}
Robillard et al. used a mixed approach of surveys and in-person interviews with professional software engineers to understand the API learning obstacles \cite{Robillard_what_makes} \cite{Robillard_a_field_study}. They found that many obstacles for API learnabilty were related to API documentation. They identified five impactful factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation. Uddin and Robillard surveyed 323 professional software engineers to understand the different ways in which a piece of documentation can be unfit for purpose \cite{g_uddin}. They identified ambiguity, incompleteness, and incorrectness as the three severest problems that lead API documentations to fail to answer the information needs of developers. Duala-Ekoko and Robillard performed a study with 20 programmers to understand the types of questions that developers have while using an unfamiliar API \cite{Duala-Ekoko:2012:AAQ:2337223.2337255}. They categorized the API related information needs into twenty generic questions that can be used to analyze the quality of API documentation. Myers et al. found that participants had limited success finding the relevant API elements to perform an API task using an enterprise API \cite{Myers_study}. Among other factors, they recommended providing code examples in the API documentation to help API client developers. Our work fits in the general space of studies of API documentation obstacles. We focus on problems related to the absence of examples within the area of REST APIs.

\subsubsection{Crowd-Sourced API Usage Examples}
Wang et al. performed an exploratory study to understand the current state of API related knowledge available on the Internet \cite{6462686}. They searched the web for usage examples related to the API of five popular Java libraries and found that on average API examples could be found for 77\% of the 4,637 APIs from the libraries. Moreover, they found that the crowd-sourcing sites accounted for 93.7\% of the usage examples compared to 6.13\% that were officially published. An evaluation of Jadite showed that developers were three times faster to complete API tasks with access to auto-referred crowd-sourced usage examples~\cite{5295283}. Nasehi et al. analyzed StackOverflow threads to understand the characteristics of good and bad code examples \cite{Nasehi_what_makes}. They found that commonly down-voted API related answers lacked code examples, explanations, and shortcoming of solutions. Treude et al. proposed a machine learning based approach to find relevant API documentation from StackOverflow using both the text and metadata found from the questions and answers~\cite{Treude:2016:AAD:2884781.2884800}. Kim et al. presented a technique to automatically augment code examples from the web using code search tools to Java APIs \cite{Kim:2009:AEJ:1747491.1747552}. Jiau et al. observed a severe inequality within the context of crowd-sourced API usage examples where most of the content were related to the popular APIs \cite{jiau}. Existing research on crowd-sourced API documentation mostly focused on finding usage examples of local APIs. In this paper, we focused on REST APIs, where the API client development programming language is agnostic to API implementation language. This distinction with local APIs makes it hard to use the techniques proposed in the aforementioned papers for finding crowd-sourced REST API usage examples.
